---
title: "🗳️ Political Data Science: A/B Testing & Causal Inference"

output: 
  html_document:
    theme: journal
    toc: true
    toc_depth: 2
    toc_float: true
    highlight: espresso
---



# 📝 Executive Summary

This project demonstrates the application of **Causal Inference** and **A/B Testing** within a political communications context. By simulating 1,500 website visitors, I evaluate how different landing page versions—a standard "Policy" page versus an "Issue-Focused" page—impact supporter signup rates.


---



## 🛠️ 1. Experimental Design & Simulation

In this randomized controlled trial (RCT), we assign users to a **Control** group (Standard Homepage) or a **Treatment** group (Issue-Focused Homepage). 

### ⚙️ Simulation Logic

* **Control Baseline:** 12% conversion rate.
* **Treatment Lift:** +5% (Total 17% expected).

```{r simulate-data}
# Set environment
set.seed(2024)
n <- 1500

# 1. Random Treatment Assignment
treatment <- rbinom(n = n, size = 1, prob = 0.5)

# 2. Potential Outcomes (Counterfactuals)
Y0 <- rbinom(n = n, size = 1, prob = 0.12) # Outcome if Control
Y1 <- rbinom(n = n, size = 1, prob = 0.17) # Outcome if Treatment

# 3. Observed Outcome (support_signup)
support_signup <- ifelse(treatment == 1, Y1, Y0)

# 4. Construct Dataframe
party_ab <- data.frame(
  visitor_id = 1:n,
  treatment = treatment,
  support_signup = support_signup
)



```

```{r}
# Load the knitr library for table formatting
library(knitr)

# Display the first 10 rows using kable
kable(head(party_ab,10), 
      caption = "Table 1: Preview of Simulated A/B Test Data",
      col.names = c("Visitor ID", "Treatment (1=Issue, 0=Control)", "Signup (1=Yes, 0=No)"),
      align = "ccc")
```



---


## 📊 2. Estimating the Average Treatment Effect (ATE)
The ATE represents the average difference in signup probability caused by the treatment.

📈 Visualizing Group Conversions

```{r}
library(ggplot2)
library(scales) # For better axis formatting

# 1. Prepare Data
mean_control <- mean(party_ab$support_signup[party_ab$treatment == 0])
mean_treated <- mean(party_ab$support_signup[party_ab$treatment == 1])
ate_hat <- mean_treated - mean_control

plot_data <- data.frame(
  Group = c("Control\n(Standard)", "Treatment\n(Issue-Focused)"),
  Rate = c(mean_control, mean_treated)
)

# 2. Generate Professional Plot
ggplot(plot_data, aes(x = Group, y = Rate, fill = Group)) +
  # Use a slightly thinner column for a modern look
  geom_col(width = 0.5, alpha = 0.85) + 
  # Add percentage labels directly above bars
  geom_text(aes(label = label_percent(accuracy = 0.1)(Rate)), 
            vjust = -0.8, 
            size = 5, 
            fontface = "bold", 
            color = "#2c3e50") +
  # Custom Professional Colors
  scale_fill_manual(values = c("Control\n(Standard)" = "#7f8c8d", 
                               "Treatment\n(Issue-Focused)" = "#2980b9")) +
  # Scale Y to give some "headroom" for labels
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.22)) +
  # Professional Titles and Labels
  labs(
    title = "Impact of Messaging on Supporter Signups",
    subtitle = paste0("The Issue-Focused homepage increased signups by ", 
                      round(ate_hat * 100, 1), 
                      " percentage points (p < 0.05)"),
    caption = "Source: Simulated A/B Test Data (n=1500)\nNote: ATE estimated via Difference-in-Means",
    x = NULL,
    y = "Conversion Rate (%)"
  ) +
  # Clean Theme Overrides
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18, color = "#2c3e50"),
    plot.subtitle = element_text(size = 13, color = "#34495e", margin = margin(b = 15)),
    plot.caption = element_text(size = 9, face = "italic", color = "gray50"),
    panel.grid.major.x = element_blank(), # Remove vertical grid lines
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "#2c3e50", face = "bold"),
    legend.position = "none"
  )

```



**Key Finding**: The Issue-Focused homepage resulted in a lift of r round(ate_hat * 100, 2) percentage points.


---


## 🔬 3. Comparative Modeling

To ensure the results are statistically sound, I compare a Linear Probability Model (LPM) against Logistic Regression.

🔹 **Linear Probability Model (OLS)**

```{r}
lm_model <- lm(support_signup ~ treatment, data = party_ab)

```

```{r echo=FALSE}
library(broom) # Helps clean 'summary' objects into data frames

# 1. Tidy the model output (estimates, std errors, p-values)
modellm_tidy <- tidy(lm_model)

# 2. Format as a professional table
kable(modellm_tidy, 
      digits = 4,
      caption = "Table 2: Regression Coefficients",
      col.names = c("Predictor", "Estimate", "Std. Error", "t-statistic", "p-value"),
      align = "lcccc")
```




🔸 **Logistic Regression (Logit)**

The Logit model handles binary outcomes by predicting log-odds, which I then convert back into probabilities.

```{r}
logit_model <- glm(support_signup ~ treatment, data = party_ab, family = binomial)

p_control <- predict(logit_model, newdata = data.frame(treatment = 0), type = "response")
p_treated <- predict(logit_model, newdata = data.frame(treatment = 1), type = "response")



```

**Logit Estimated ATE**:" `r round(p_treated - p_control, 4) `



---

## 👥 4. Demographic Heterogeneity
I analyze whether the treatment effect changes based on visitor Age or Political Interest using interaction terms.

```{r}
# Simulate additional variables
party_ab$age <- round(pmin(pmax(rnorm(n, 40, 12), 18), 75))
party_ab$high_interest <- rbinom(n, 1, 0.4)

```

### Interaction Model

**Formula** : Y ~ Treatment * Covariate

```{r}
lm_interaction <- lm(support_signup ~ treatment * high_interest + treatment * age, data = party_ab)

```

```{r echo=FALSE}
library(broom) # Helps clean 'summary' objects into data frames

# 1. Tidy the model output (estimates, std errors, p-values)
model_tidy <- tidy(lm_interaction)

# 2. Format as a professional table
kable(model_tidy, 
      digits = 4,
      caption = "Table 3: Regression Coefficients with Interaction Terms",
      col.names = c("Predictor", "Estimate (Effect Size)", "Std. Error", "t-statistic", "p-value"),
      align = "lcccc")
```



**Observation**: If the treatment:age coefficient is positive, it suggests that the Issue-Focused page becomes more effective as the audience gets older.


---

## ⚖️ 5. Ethical Implications

In political data science, optimization must be balanced with ethics:

**Transparency**: Unlike commercial A/B tests, political messaging influences democratic choice.

**Micro-targeting**: Using demographic interactions to serve specific messages can lead to societal polarization.

**Data Privacy**: The collection of "Interest Level" creates sensitive profiles that require high security.



---

## 🏁 Conclusion
The data confirms that an Issue-Focused landing page significantly increases conversion for this fictitious party. Moving forward, the party should consider a rollout of Version B while maintaining a commitment to factual accuracy and ethical voter engagement.